---
title: "Data processing"
author: "Richard J Beck"
date: "3/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="C:/Users/4473331/Documents/projects/006_ploidyModeling/06_ploidy_glucose/glucose_wgd/")
```

```{r}
library(ggplot2)
source("Rscripts/utils.R")

```

```{r}
fit_dat <- list()
plot_dat <- list()
```

Processing data from first experiment.
One issue to consider is how we get the standard deviation of the glucose measurements. I think we need to do something more sophisticated than simply take the std.dev of the measurements? With the current approach, I think we are effectively weighting the glucose data from experiment 1 more heavily than that from experiment 2. 

```{r}

x <- readRDS("data/raw/220823_ilastik_counts.Rds")
x$day <- (x$frame*8+8)/24

bx <- aggregate(list(ncells=x$ncells),by=list(label=x$label,
                                                day=x$day,
                                                ploidy=x$ploidy,
                                              glucose=x$glucose),mean)
bx$sd <- aggregate(list(ncells=x$ncells),by=list(label=x$label,
                                                day=x$day,
                                                ploidy=x$ploidy,
                                              glucose=x$glucose),sd)$ncells

fit_dat$bx1 <- bx
bx$id <- "exp 1"
plot_dat$bx <- bx

xg <- read.csv("data/raw/220823_Glucose consumption_Sum159_2N_4N.csv")
colnames(xg)[2:6] <- c(0.1,0.5,1,5,25)
xg <- reshape2::melt(xg,id.vars=c("rep","day","ploidy"))
xg$value <- as.numeric(xg$value)

xg$value <- xg$value/18.018

gx <- aggregate(list(G=xg$value),by=list(day=xg$day,
                                                ploidy=xg$ploidy,
                                                glucose=xg$variable),mean)
gx$sd <- aggregate(list(G=xg$value),by=list(day=xg$day,
                                                ploidy=xg$ploidy,
                                                glucose=xg$variable),sd)$G

## some points have zero std.dev. 
opt <- lm(sd~G,data=gx)

gx$G[is.na(gx$G)] <- 0.2576315 # mean(runif(10000,0,0.5))
gx$sd[is.na(gx$sd)] <- 0.14892716 # sd(runif(10000,0,0.5))
gx$sd[gx$sd==0] <- predict(opt,gx)[gx$sd==0]
gx$glucose <- as.character(gx$glucose)
fit_dat$gx1 <- gx

gx$Gmin <- gx$G-gx$sd
gx$Gmax <- gx$G+gx$sd

gx <- gx[,!colnames(gx)=="sd"]
gx$id <- "exp 1"
plot_dat$gx <- gx

p <- ggplot(bx,aes(x=day,y=ncells,color=label))+
  facet_grid(cols=vars(ploidy),rows=vars(glucose),scales="free")+
  geom_point()+
  geom_errorbar(aes(ymin=ncells-sd,ymax=ncells+sd))
p

p <- ggplot(gx,aes(x=day,y=G))+
  facet_grid(cols=vars(ploidy),rows=vars(glucose),scales="free")+
  geom_point()+
  geom_errorbar(aes(ymin=Gmin,ymax=Gmax))
p

```

Processing data from experiment 2.

First perform the glucose calibration and save the relevant parameters and functions:

```{r}


glu_prob <- function(glu,lum,gpar,aslog=F){
  Elum <- gpar["a"]*glu + gpar["b"]
  dnorm(log(lum)-log(Elum),sd=gpar["sd"],log = aslog)
}

glu_quantiles <- function(lum,q=c(0.95,0.5,0.05),gpar){
  (exp(log(lum)-qnorm(p=q,sd=gpar["sd"]))-gpar["b"])/gpar["a"]
}

fit_log <- function(pars,dat){
  sqrt(mean((log(dat$G*abs(pars[1])+abs(pars[2]))-log(dat$value))^2))
}
x <- read.csv("data/raw/221112_glucose_consumption.csv")
x <- reshape2::melt(x,id.vars=c("G","Dilution","day","ploidy"))
xcalib <- x[x$day==0&x$G>0,]
xcalib$G <- xcalib$G/xcalib$Dilution

p0 <- c(150000,20000)*1000

opt <- optim(p0,fit_log,dat=xcalib)
pars <- abs(opt$par)

pred_F <- xcalib$G*abs(pars[1])+abs(pars[2])

sdlog <- sd(log(pred_F)-log(xcalib$value))

glu_pars <- c(pars,sdlog)
names(glu_pars) <- c("a","b","sd")

glucose_struct <- list(gpar=glu_pars,pg=glu_prob,qp=glu_quantiles)

glucose_struct$qp(30000,gpar=glucose_struct$gpar)
saveRDS(glucose_struct,"data/fitted_parameters/glucose_struct.Rds")

```

Note that there is a difference in survivorship between experiments 1 & 2. This does not appear to be related to errors in cell counting, could there be differences in the media instead??

```{r}

x <- readRDS("data/raw/221112_ilastik_counts.Rds")
x$day <- (x$frame*8+8)/24

bx <- aggregate(list(ncells=x$ncells),by=list(label=x$label,
                                                day=x$day,
                                                ploidy=x$ploidy,
                                              glucose=x$glucose),mean)
bx$sd <- aggregate(list(ncells=x$ncells),by=list(label=x$label,
                                                day=x$day,
                                                ploidy=x$ploidy,
                                              glucose=x$glucose),sd)$ncells
fit_dat$bx2 <- bx

bx$id <- "exp 2"
plot_dat$bx <- rbind(plot_dat$bx,bx)
p <- ggplot(bx,aes(x=day,y=ncells,color=label))+
  facet_grid(cols=vars(ploidy),rows=vars(glucose),scales="free")+
  geom_point()+
  geom_errorbar(aes(ymin=ncells-sd,ymax=ncells+sd))
p

gx <- read.csv("data/raw/221112_glucose_consumption.csv")
gs <- readRDS("data/fitted_parameters/glucose_struct.Rds")

gx <- gx[!gx$Dilution==1,]
tmp <- gx[gx$ploidy=="NaN" ]
tmp$ploidy <- "2N"
gx$ploidy[gx$ploidy=="NaN"] <- "4N"
gx <- rbind(tmp,gx)

gx <- reshape2::melt(gx,id.vars=c("G","Dilution","day","ploidy"))

gx <- aggregate(list(lum=gx$value),by=list(glucose=gx$G,dilution=gx$Dilution,
                                           day=gx$day,ploidy=gx$ploidy),mean)
gx <- gx[gx$glucose<5,]
fit_dat$gx2 <- gx


gxp <- data.frame(do.call(rbind,lapply(gx$lum,gs$qp,gpar=gs$gpar)))
gxp <- gxp*gx$dilution
colnames(gxp) <- c("Gmin","G","Gmax")
gx <- cbind(gx[,c("day","ploidy","glucose")],gxp)
gx$id <- "exp 2"
plot_dat$gx <- rbind(plot_dat$gx,gx)

p <- ggplot(gx,aes(x=day,y=G))+
  facet_grid(cols=vars(ploidy),rows=vars(glucose),scales="free")+
  geom_point()+
  geom_errorbar(aes(ymin=Gmin,ymax=Gmax))
p

saveRDS(plot_dat,"data/processed/plot_dat.Rds")
saveRDS(fit_dat,"data/fitting/fit_dat.Rds")

```
```{r}


p <- ggplot(plot_dat$gx,aes(x=day,y=G,color=id))+
  facet_grid(cols=vars(ploidy),rows=vars(glucose),scales="free")+
  geom_point()+
  geom_errorbar(aes(ymin=Gmin,ymax=Gmax))
p

p <- ggplot(plot_dat$bx,aes(x=day,y=ncells,color=interaction(label,id)))+
  facet_grid(cols=vars(ploidy),rows=vars(glucose),scales="free")+
  geom_point()+
  geom_errorbar(aes(ymin=ncells-sd,ymax=ncells+sd))
p

```